---
title: "fhidata"
output: rmarkdown::html_vignette
author: "Richard Aubrey White"
date: "2021-02-18"
vignette: >
  %\VignetteIndexEntry{fhidata}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


`fhidata` provides structural data for Norway.

Please see https://folkehelseinstituttet.github.io/fhidata/reference/index.html for a quick overview of all available datasets and functions.

You can find out what version of `fhidata` you have installed by running the following commands:

```{r setup}
library(fhidata)
library(data.table)
```


## granularity_geo

The underlying foundation of the `fhidata` package is the strict categorization of `location_code` into `granularity_geo`. The concept is that a person should (generally) have data for all values of `location_code` for each `granularity_geo` that they are interested in.

```{r}
# All the available values of granularity_geo
unique(fhidata::norway_locations_names()$granularity_geo)
```

So if a person is interested in doing analyses for `granularity_geo=="county"` then it is expected that a person has data for all of the following values of `location_code`

```{r}
# All the available values location_code for granularity_geo=="county"
fhidata::norway_locations_names()[granularity_geo=="county"]$location_code
```

This is the reason why there are three "similar" values for

- "county": `county`, `notmainlandcounty`, `missingcounty`.
- "municip": `municip`, `notmainlandmunicip`, `missingmunicip`.

i.e. A person may want to use county-level data, but not want to include counties/municipalities that are not on mainland Norway.

## Accessing Norwegian data via helper functions

It is recommended to access Norwegian population and location datasets via helper functions.

### Norwegian locations

Redistricting is a common feature of the present Norwegian political climate. It is therefore important that our datasets can handle this. It is therefore recommended that a person set their desired year of redistricting (currently only 2020 is available) before using Norwegian data from this package. It is also possible to customize the desired border on a per-function basis using the `border` argument.

```{r}
fhidata::set_config(border = 2020)
```

When using historical data, it is important to map the old/original `location_code` to the current/desired `location_code`.

```{r}
fhidata::norway_locations_redistricting()
```

It is also important to understand the hierarchy of Norwegian locations, so that you can aggregate your own data up to different levels of `granularity_geo`.

```{r}
fhidata::norway_locations_hierarchy(from = "wardoslo", to="municip")
fhidata::norway_locations_hierarchy(from = "municip", to="baregion")
fhidata::norway_locations_hierarchy(from = "county", to="faregion")
```

When creating a skeleton of a dataset, it is important to know all of the values of `location_code` within your desired `granularity_geo`.

```{r}
fhidata::norway_locations_names()[granularity_geo %in% c("county", "notmainlandcounty","missingcounty")]$location_code
```

Finally, for publication purposes it is important to have a) names of the locations, and b) the correct ordering of the locations.

```{r}
fhidata::norway_locations_names()
```


